---
title: "ESM 262 Assignment 3"
author: "Cameryn Brock"
date: "2/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

### (Simplified) Function for Reaching Functional Extinction

*If a species declines by x% per year, how long until the population will reach functional extinction?* 

```{r}
source("R/time_extinc.R")
```

Example: 

* Species name: Fire dragon
* Initial population size: 10,000 individuals
* Population size at function extinction: 50 individuals
* Maximum time to project out to: 1,000 years


```{r}
# set constants (for this example - not universal)
n_e = 50
t_max = 1000
t_0 = 0

# If declining 10% per year 
time_extinc_fun(
  t = t_0,
  n_t = 10000,
  p_d = (-0.1),
  n_e = n_e,
  t_max = t_max
)
```
If declining 10% per year, will reach functional extinction in 51 years.


```{r}
# if declining 1% per year
time_extinc_fun(
  t = t_0,
  n_t = 10000,
  p_d = (-0.01),
  n_e = n_e,
  t_max = t_max
)
```
If declining 1% per year, functional extinction in 528 years.

### Make df with `sapply`

```{r}
pct_d = seq(from = -0.005, to = -0.6, by = -0.005)

extinc_times <- pct_d %>% 
  sapply(time_extinc_fun,
          t = t_0, n_t = 5000,
          n_e = n_e, t_max = t_max)

extinc_times_df <- data.frame(pct_decline = pct_d, 
                              extinc_time = extinc_times)
```

### Add level of concern with `case_when`

```{r}
extinc_concern <- extinc_times_df %>% 
  mutate(concern = case_when(
    extinc_time >= 500 ~ "Low",
    extinc_time < 500 & extinc_time > 250 ~ "Moderate",
    extinc_time < 250 & extinc_time > 50 ~ "High",
    extinc_time <= 50 ~ "Very High"))
```

### Visualize Results

```{r}
extinc_concern$concern <- factor(extinc_concern$concern, 
                                 levels = c("Low", "Moderate", 
                                            "High", "Very High"))

ggplot(extinc_concern, aes(x = pct_decline,
                           y = extinc_time)) + 
  geom_point(aes(color = concern),
             alpha = 0.9) + 
  scale_color_manual(values = c("#1481BA", "#68A691", "#f19953", "#9b1d20")) +
  scale_x_continuous(trans = "reverse") +
  theme_minimal() +
  theme(legend.position = "top") +
  labs(color = "Level of concern",
       x = "Annual decline (%)",
       y = "Time to functional extinction (years)")
```






